---
title: "gridcell-attempt"
format: html
editor: visual
---
#libraries

```{r}

#| message: false 
#| warning: false 

library(terra)
library(rnaturalearth)
library(dplyr)
library(viridisLite)
library(fixest)
library(tidyverse) 
library(sf) 
library(here) 
library(stars) 
library(tmap) 
library(viridis) 
library(readr)
library(dplyr)
library(rnaturalearth)
library(rnaturalearthdata)
library(lubridate)
library(terra)
library(exactextractr)
library(readr)
library(haven)
library(sf)
library(units)
library(car)
```

#reading it in

```{r}
#| message: false 
#| warning: false 
ab <- read_sav("ABdata/R9Geo.sav")
glbl <- rast("Popdata/Global.tif")
acled <- sf::st_read(here::here("ACLEDnew.csv")) #its giving a warning that it's not spatial, thats ok i know
names(acled)[grep("lat|lon|long", names(acled), ignore.case = TRUE)]

lat_col <- "latitude"
lon_col <- "longitude"
acled <- acled %>%
  mutate(
    !!lat_col := as.numeric(.data[[lat_col]]),
    !!lon_col := as.numeric(.data[[lon_col]])
  ) %>%
  filter(!is.na(.data[[lat_col]]), !is.na(.data[[lon_col]]),
         between(.data[[lat_col]], -90, 90),
         between(.data[[lon_col]], -180, 180))
acled_sf <- st_as_sf(acled, coords = c(lon_col, lat_col), crs = 4326, remove = FALSE)
```
####### Cropping to WA
```{r}
#| message: false 
#| warning: false 
africa <- ne_countries(continent = "Africa", scale = "medium", returnclass = "sf")
# 3. Filter for your West African countries
wa_countries <- africa %>%
  dplyr::filter(admin %in% c(
    "Benin", "Burkina Faso", "Cabo Verde", "Ivory Coast",
    "Gambia", "Ghana", "Guinea", "Guinea Bissau", "Liberia", "Mali", "Mauritania", "Niger",
    "Nigeria", "Senegal", "Sierra Leone", "Togo"
  ))
wa_v <- vect(wa_countries)

# 5. Crop and mask the raster
glbl_wa <- crop(glbl, wa_v)
glbl_wa <- mask(glbl_wa, wa_v)

# 6. Plot result
plot(glbl_wa)
```


####### Cropping to WA

```{r}
#| message: false 
#| warning: false 
africa <- ne_countries(continent = "Africa", scale = "medium", returnclass = "sf")
# 3. Filter for your West African countries
wa_countries <- africa %>%
  dplyr::filter(admin %in% c(
    "Benin", "Burkina Faso", "Cabo Verde", "Ivory Coast",
    "Gambia", "Ghana", "Guinea", "Guinea Bissau", "Liberia", "Mali", "Mauritania", "Niger",
    "Nigeria", "Senegal", "Sierra Leone", "Togo"
  ))
wa_v <- vect(wa_countries)

# 5. Crop and mask the raster
glbl_wa <- crop(glbl, wa_v)
glbl_wa <- mask(glbl_wa, wa_v)

# 6. Plot result
plot(glbl_wa)
```

#configuring for regressions

```{r}
#| message: false 
#| warning: false 
guess_coord <- function(df){
  nms <- names(df)
  lat_guess <- nms[grep("EA_GPS_LA", nms, ignore.case = TRUE)][1]
  lon_guess <- nms[grep("EA_GPS_LO", nms, ignore.case = TRUE)][1]
  stopifnot(!is.na(lon_guess), !is.na(lat_guess))
  list(lon = lon_guess, lat = lat_guess)
}
xy <- guess_coord(ab)
#names(ab)

ab_sf <- ab %>%
  mutate(
    !!xy$lon := as.numeric(.data[[xy$lon]]),
    !!xy$lat := as.numeric(.data[[xy$lat]])
  ) %>%
  filter(
    between(.data[[xy$lat]], -90, 90),
    between(.data[[xy$lon]], -180, 180)
  ) %>%
  st_as_sf(coords = c(xy$lon, xy$lat), crs = 4326, remove = FALSE)

#n  = 53444
```

population density

```{r}
#| message: false 
#| warning: false 
# --------------------------------------
# 3) Population density at respondent
#     (convert counts -> per km²)
# --------------------------------------
# If Global.tif is population *counts per cell*, convert to density:
cell_km2 <- terra::cellSize(glbl, unit = "km")
pop_dens <- glbl / cell_km2   # persons per km²

# Extract density for each AB point
# terra::extract() works with sf; returns a data.frame with ID + values
ab_sf$pop_density <- terra::extract(pop_dens, ab_sf)[, 2]

# Optional: cap huge outliers & log-scale for stability
ab_sf <- ab_sf %>%
  mutate(pop_density_log = log1p(pmin(pop_density, quantile(pop_density, 0.99, na.rm = TRUE))))

dist_25km <- set_units(25, km)

within_25km <- st_is_within_distance(
  ab_sf,
  acled_sf,
  dist = dist_25km
)

within_25km <- st_is_within_distance(ab_sf, acled_sf, dist = set_units(25, "km"))
conflict_counts <- lengths(within_25km)

ab_sf <- ab_sf %>%
  mutate(
    conflict_count     = conflict_counts,
    conflict_presence  = conflict_counts > 0
  )

```

actually working with the questions

```{r}
q37j_raw <- ab$Q37J

# Drop special missing codes
bad_codes <- c(8, 9, -1)
q37j_raw[q37j_raw %in% bad_codes] <- NA

# Now this is already 0–3
trust_trad_leaders <- q37j_raw

q37j_raw <- ab$Q37J
q37j_raw[q37j_raw %in% c(8, 9, -1)] <- NA
#trust_trad_leaders <- make_trust_0_3(q37j_raw)

```

```{r}
make_trust_0_3 <- function(x) {
  # convert to numeric if labelled/factor
  x_num <- suppressWarnings(as.numeric(x))

  # missing / special codes to NA
  x_num[x_num %in% c(8, 9, 98, 99, -1, -2)] <- NA

  # if everything is NA, just return NAs
  if (all(is.na(x_num))) return(x_num)

  rng <- range(x_num, na.rm = TRUE)

  # 0–3: already on target scale
  if (rng[1] >= 0 && rng[2] <= 3) {
    out <- x_num

  # 1–4: shift down by 1
  } else if (rng[1] >= 1 && rng[2] <= 4) {
    out <- x_num - 1

  # anything else: leave as is, then clean
  } else {
    out <- x_num
  }

  # enforce 0–3 support
  out[!is.na(out) & !(out %in% 0:3)] <- NA_real_

  out
}

```

Mkaing an index

```{r}
#| message: false 
#| warning: false 
#creating an index
trust_candidates <- c("Q37J", "Q37I", 'Q37F', 'Q37C', 'Q37D')
#trust_corrpution <- c("Q37J", "Q37I", 'Q37F', 'Q37C', 'Q37B', 'Q37D', 'Q37K', 'Q37I', "Q38D", "Q38E", "Q38F", "Q38G", "Q38H")
trust_cols <- trust_candidates[trust_candidates %in% names(ab_sf)]
if (length(trust_cols) == 0) {
  stop("Couldn't find any trust variables. Add their names to 'trust_candidates'.")
}

ab_sf <- ab_sf %>%
  mutate(
    across(
      all_of(trust_cols),
      make_trust_0_3,
      .names = "{.col}_0_3"
    )
  )
#trust_corr <- trust_corrpution[trust_corrpution %in% names(ab_sf)]
#if (length(trust_corr) == 0) {
#  stop("Couldn't find any trust variables. Add their names to 'trust_candidates'.")

ab_sf <- ab_sf %>%
  mutate(
    across(
      all_of(trust_cols),
      make_trust_0_3,
      .names = "{.col}_0_3"
    )
  )


trust_index_cols <- paste0(trust_cols, "_0_3")

ab_sf <- ab_sf %>%
  rowwise() %>%
  mutate(
    trust_index = if_else(
      # require at least 2 non-missing items (you can change this)
      sum(!is.na(c_across(all_of(trust_index_cols)))) >= 2,
      mean(c_across(all_of(trust_index_cols)), na.rm = TRUE),
      NA_real_
    )
  ) %>%
  ungroup()


# controls
ctrls <- intersect(c("Q1", "Q2", "URBRUR", "Q100", "Q94", "Q95"), names(ab_sf))
# Country/region FEs if present
fe_country <- intersect(c("country", "country_name", "COUNTRY"), names(ab_sf))[1]
fe_region  <- intersect(c("region", "region_name", "ADM1NAME"), names(ab_sf))[1]


```

```{r}
#| message: false 
#| warning: false 
density_var <- if ("pop_density" %in% names(ab_sf)) {
  "pop_density"
} else if ("pop_density_log" %in% names(ab_sf)) {
  "pop_density_log"
} else {
  stop("No population density variable found (pop_density or pop_density_log).")
}


# main regressors: trust + density + other controls
rhs_vars <- c("trust_index", density_var, ctrls)

# collapse into a single RHS string: e.g. "trust_index + pop_density + Q1 + Q2 + URBRUR + ..."
rhs_str <- paste(rhs_vars, collapse = " + ")

# Start with the basic part: DV ~ regressors
base_str <- paste0("conflict_presence ~ ", rhs_str)

# Build FE part depending on what exists
fe_part <- NULL

if (!is.na(fe_country) && !is.na(fe_region)) {
  fe_part <- paste(fe_country, fe_region, sep = " + ")
} else if (!is.na(fe_region)) {
  fe_part <- fe_region
} else if (!is.na(fe_country)) {
  fe_part <- fe_country
}

# Combine into final formula string
if (!is.null(fe_part)) {
  # fixest syntax: y ~ x1 + x2 | fe1 + fe2
  f_str <- paste(base_str, "|", fe_part)
} else {
  # no FEs available
  f_str <- base_str
}

f_logit <- as.formula(f_str)
f_logit



```
#models
```{r}
```{r}
#| message: false 
#| warning: false 
library(fixest)

m_logit <- feglm(
  fml    = f_logit,
  data   = ab_sf,
  family = "binomial"
)

summary(m_logit)
library(stargazer)

ct <- summary(m_logit)$coeftable
# columns: "Estimate" "Std. Error" "z value" "Pr(>|z|)"

coefs <- ct[, "Estimate"]
ses   <- ct[, "Std. Error"]
pvals <- ct[, "Pr(>|z|)"]

# make sure output dir exists
if (!dir.exists("Outputs")) dir.create("Outputs")

# Drop geometry so glm doesn't choke on the sf column
dat <- st_drop_geometry(ab_sf)

# Use the same FE variable you already detected (fe_country)
# (If fe_country is NA, this will error—good, because it means you don't have a country column.)
stopifnot(!is.na(fe_country))

m1 <- glm(
  reformulate(
    c("trust_index", density_var, ctrls, paste0("factor(", fe_country, ")")),
    response = "conflict_presence"
  ),
  data = dat,
  family = binomial(link = "logit")
)


stargazer(
  m1,
  type = "latex",
  title = "Logit estimates of conflict presence",
  dep.var.labels = "Conflict presence (binary)",
  omit = fe_country,  # omits all factor(COUNTRY) dummy rows
  add.lines = list(
    c("Country fixed effects", "Yes")
  ),
  digits = 3,
  no.space = TRUE
)
```
#trust by gridcell
```{r}
library(sf)
library(dplyr)

# Choose an equal-area CRS so grid cell size in meters is meaningful
crs_ea <- 6933  # World Cylindrical Equal Area (works fine for WA-scale summaries)

# Transform WA polygon + AB points to equal-area
wa_sf_ea  <- st_transform(wa_countries, crs_ea)
ab_sf_ea  <- st_transform(ab_sf, crs_ea)

# (optional but recommended) drop empty geometries / invalid
wa_sf_ea  <- st_make_valid(wa_sf_ea)

# Grid size in meters
cell_km <- 75
cell_m  <- cell_km * 1000

# Build grid, then clip to WA (so you don't keep ocean cells)
grid_ea <- st_make_grid(
  wa_sf_ea,
  cellsize = cell_m,
  square   = TRUE
) |>
  st_as_sf() |>
  mutate(cell_id = row_number())

# Keep only cells that intersect West Africa
wa_union <- st_union(wa_sf_ea)
grid_wa  <- grid_ea[st_intersects(grid_ea, wa_union, sparse = FALSE), ]


# Join cell_id onto each AB point (left join: keep all AB points)
ab_with_cell <- st_join(
  ab_sf_ea,
  grid_wa[, c("cell_id")],
  join = st_within,
  left = TRUE
)
ab_with_cell <- ab_with_cell |>
  mutate(cell_id = cell_id.y) |>
  select(-cell_id.x, -cell_id.y)

trust_by_cell <- ab_with_cell |>
  st_drop_geometry() |>
  group_by(cell_id) |>
  summarise(
    n_resp          = sum(!is.na(trust_index)),
    trust_mean      = mean(trust_index, na.rm = TRUE),
    trust_sd        = sd(trust_index, na.rm = TRUE),
    popdens_mean    = mean(pop_density, na.rm = TRUE),
    conflict_any_sh = mean(conflict_presence, na.rm = TRUE)  # share of respondents in cell w/ conflict within 25km
  ) |>
  ungroup()

# Attach summaries back onto the grid polygons
grid_trust <- grid_wa |>
  left_join(trust_by_cell, by = "cell_id")
c(
  x_na = mean(is.na(ab_with_cell$cell_id.x)),
  y_na = mean(is.na(ab_with_cell$cell_id.y))
)

trust_by_cell <- ab_with_cell |>
  st_drop_geometry() |>
  group_by(cell_id) |>
  summarise(
    n_resp          = sum(!is.na(trust_index)),
    trust_mean      = mean(trust_index, na.rm = TRUE),
    trust_sd        = sd(trust_index, na.rm = TRUE),
    popdens_mean    = mean(pop_density, na.rm = TRUE),
    conflict_any_sh = mean(conflict_presence, na.rm = TRUE)  # share of respondents in cell w/ conflict within 25km
  ) |>
  ungroup()

# Attach summaries back onto the grid polygons
grid_trust <- grid_wa |>
  left_join(trust_by_cell, by = "cell_id")
```
##map
```{r}
library(tmap)

tm_shape(grid_trust) +
  tm_polygons(
    col = "trust_mean",
    style = "cont",
    title = paste0("Mean trust index (", cell_km, "km cells)")
  ) +
  tm_layout(frame = FALSE)
```
#conflict by grid cell
```{r}
library(lubridate)
library(sf)
library(dplyr)

# If your date col isn't event_date, change it here:
date_col <- "event_date"

acled_2020_2024 <- acled_sf %>%
  mutate(event_date = as.Date(.data[[date_col]])) %>%   # if already Date, this is fine
  filter(event_date >= as.Date("2020-01-01"),
         event_date <= as.Date("2024-12-31"))

# project to same CRS as grid
acled_ea <- st_transform(acled_2020_2024, crs_ea)

acled_with_cell <- st_join(
  acled_ea |> dplyr::select(-any_of("cell_id")),              # drop any existing cell_id
  grid_wa  |> dplyr::select(cell_id) |> rename(grid_cell = cell_id),
  join = st_within,
  left = FALSE
)

conflict_by_cell <- acled_with_cell %>%
  st_drop_geometry() %>%
  count(grid_cell, name = "conflict_events_2020_2024") %>%
  mutate(conflict_avg_annual_2020_2024 = conflict_events_2020_2024 / 5)

grid_conflict <- grid_wa %>%
  rename(grid_cell = cell_id) %>%
  left_join(conflict_by_cell, by = "grid_cell") %>%
  mutate(
    conflict_events_2020_2024 = tidyr::replace_na(conflict_events_2020_2024, 0),
    conflict_avg_annual_2020_2024 = tidyr::replace_na(conflict_avg_annual_2020_2024, 0)
  )
```
##map
```{r}
library(tmap)

tm_shape(grid_conflict) +
  tm_polygons(col = "conflict_avg_annual_2020_2024",
              style = "cont",
              title = "Avg annual conflict events (2020–2024)") +
  tm_layout(frame = FALSE)

```

#trustxconflict
```{r}
# Make trust grid use the same id name as conflict grid
grid_trust2 <- grid_trust %>%
  rename(grid_cell = cell_id)

# Join trust onto the conflict grid (or vice versa)
grid_analysis <- grid_conflict %>%
  left_join(
    grid_trust2 %>% st_drop_geometry() %>%
      select(grid_cell, n_resp, trust_mean, trust_sd, popdens_mean),
    by = "grid_cell"
  )

# Optional but smart: drop cells with too few AB respondents
min_n <- 0
grid_analysis <- grid_analysis %>%
  mutate(trust_mean = ifelse(n_resp >= min_n, trust_mean, NA_real_))

library(MASS)

df <- grid_analysis %>%
  st_drop_geometry() %>%
  filter(!is.na(trust_mean))

m_nb <- MASS::glm.nb(
  conflict_events_2020_2024 ~ trust_mean + log1p(popdens_mean),
  data = df
)

summary(m_nb)

df <- grid_analysis %>%
  mutate(cell_area_km2 = as.numeric(st_area(geometry)) / 1e6) %>%
  st_drop_geometry() %>%
  filter(!is.na(trust_mean))

m_nb_area <- MASS::glm.nb(
  conflict_events_2020_2024 ~ trust_mean + log1p(popdens_mean) + log(cell_area_km2),
  data = df
)

summary(m_nb_area)
```