---
title: "Gridcells"
author: "Maeve Li Oak"
format: html
editor: visual
---

#packages
```{r}
#reading libraries
library(tidyverse) 
library(sf) 
library(here) 
library(stars) 
library(tmap) 
library(viridis) 
library(readr)
library(dplyr)
library(rnaturalearth)
library(rnaturalearthdata)
library(lubridate)
library(terra)
library(exactextractr)
library(readr)
library(fixest)
```

#reading in data
```{r}
#population-this one is big and takes forever...just know that in your spirit as you move forward
pop_rast <- terra::rast(here::here("popdata", "Global.tif"))
pop_ea <- terra::project(pop_rast, paste0("EPSG:", crs_ea))
#afrobarometer
r6 <- sf::st_read(here::here("requestedAB", "Maeve_r6.csv")) 
r7 <- haven::read_sav(here::here("requestedAB", "Maeve_r7.sav"))
r8 <- haven::read_sav(here::here("requestedAB", "Maeve_r8.sav"))
r9 <- haven::read_sav(here::here("ABdata/R9Geo.sav"))

#acled
acled <- sf::st_read(here::here("ACLEDnew.csv")) 
names(acled)[grep("lat|lon|long", names(acled), ignore.case = TRUE)]

lat_col <- "latitude"
lon_col <- "longitude"

# Ensure numeric & sane values
acled <- acled %>%
  mutate(
    !!lat_col := as.numeric(.data[[lat_col]]),
    !!lon_col := as.numeric(.data[[lon_col]])
  ) %>%
  filter(!is.na(.data[[lat_col]]), !is.na(.data[[lon_col]]),
         between(.data[[lat_col]], -90, 90),
         between(.data[[lon_col]], -180, 180))

acled_sf <- st_as_sf(acled, coords = c(lon_col, lat_col), crs = 4326, remove = FALSE)
```


```{r}
#making a basemap----
west_af <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") |>
  filter(subregion == "Western Africa") |>
  st_transform(st_crs(acled_sf)) |>
  st_make_valid()

acled_wa <- acled_sf |>
  st_make_valid() |>
  st_intersection(st_union(st_geometry(west_af)))

tmap_mode("view")

```
building grid
```{r}
crs_ea <- 6933
cell_km <- 75
cell_m  <- cell_km * 1000

west_af_ea <- st_transform(west_af, crs_ea)
wa_union_ea <- st_union(st_geometry(west_af_ea))

grid_ea <- st_make_grid(
  wa_union_ea,
  cellsize = cell_m,
  square = TRUE
) %>%
  st_as_sf() %>%
  mutate(cell_id = row_number()) %>%
  st_intersection(wa_union_ea)

grid <- st_transform(grid_ea, 4326)
```

#helper function
```{r}
recode_ab_trust <- function(x) {
  x <- as.numeric(x)
  x[x %in% c(-1, 8, 9, 98, 99)] <- NA
  return(x)
}
```

#round 6
```{r}
prep_r6 <- function(df) {
  df %>%
    mutate(
      round = 6,
      survey_year = 2016,
      trust_president    = recode_ab_trust(q52a),
      trust_parliament   = recode_ab_trust(q52b),
      trust_electoral    = recode_ab_trust(q52c),
      trust_tax          = recode_ab_trust(q52d),
      trust_localgov     = recode_ab_trust(q52e),
      trust_rulingparty  = recode_ab_trust(q52f),
      trust_opposition   = recode_ab_trust(q52g),
      trust_police       = recode_ab_trust(q52h),
      trust_army         = recode_ab_trust(q52i),
      trust_courts       = recode_ab_trust(q52j),
      trust_traditional  = recode_ab_trust(q52k),
      trust_religious    = recode_ab_trust(q52l),
      age = as.numeric(q1),
      gender = as.numeric(q101),
      ethnicity = as.character(q87),
      ethnic_unfair = as.numeric(q88a),
      religion = as.character(q98a),
      lon = as.numeric(longitude),
      lat = as.numeric(latitude)
    ) %>%
    select(round, survey_year, lon, lat, age, gender, ethnicity, ethnic_unfair, religion,
           starts_with("trust_"))
}
```
#round 7
```{r}
prep_r7 <- function(df) {
  df %>%
    mutate(
      round = 7,
      survey_year = 2020,   # adjust if needed
      trust_president    = recode_ab_trust(Q43A),
      trust_parliament   = recode_ab_trust(Q43B),
      trust_electoral    = recode_ab_trust(Q43C),
      trust_localgov     = recode_ab_trust(Q43D),
      trust_rulingparty  = recode_ab_trust(Q43E),
      trust_opposition   = recode_ab_trust(Q43F),
      trust_police       = recode_ab_trust(Q43G),
      trust_army         = recode_ab_trust(Q43H),
      trust_courts       = recode_ab_trust(Q43I),
      trust_traditional  = recode_ab_trust(Q43J),
      trust_religious    = recode_ab_trust(Q43K),
      age = as.numeric(Q1),
      gender = as.numeric(Q101),
      ethnicity = as.character(Q84),
      ethnic_unfair = as.numeric(Q85A),
      religion = as.character(Q98),
      lon = as.numeric(EA_GPS_LO),
      lat = as.numeric(EA_GPS_LA)
    ) %>%
    select(round, survey_year, lon, lat, age, gender, ethnicity, ethnic_unfair, religion,
           starts_with("trust_"))
}
```

#round 8
```{r}
prep_r8 <- function(df) {
  df %>%
    mutate(
      round = 8,
      survey_year = 2022,   # adjust if needed
      trust_president    = recode_ab_trust(Q41A),
      trust_parliament   = recode_ab_trust(Q41B),
      trust_electoral    = recode_ab_trust(Q41C),
      trust_localgov     = recode_ab_trust(Q41D),
      trust_rulingparty  = recode_ab_trust(Q41E),
      trust_opposition   = recode_ab_trust(Q41F),
      trust_police       = recode_ab_trust(Q41G),
      trust_army         = recode_ab_trust(Q41H),
      trust_courts       = recode_ab_trust(Q41I),
      trust_tax          = recode_ab_trust(Q41J),
      trust_traditional  = recode_ab_trust(Q41K),
      trust_religious    = recode_ab_trust(Q41L),
      age = as.numeric(Q1),
      gender = as.numeric(Q101),
      ethnicity = as.character(Q81),
      ethnic_unfair = as.numeric(Q82A),
      religion = as.character(Q98A),
      lon = as.numeric(EA_GPS_LO),
      lat = as.numeric(EA_GPS_LA)
    ) %>%
    select(round, survey_year, lon, lat, age, gender, ethnicity, ethnic_unfair, religion,
           starts_with("trust_"))
}
```

#round 9
```{r}
prep_r9 <- function(df) {
  df %>%
    mutate(
      round = 9,
      survey_year = 2023,
      trust_president    = recode_ab_trust(Q37A),
      trust_parliament   = recode_ab_trust(Q37B),
      trust_electoral    = recode_ab_trust(Q37C),
      trust_localgov     = recode_ab_trust(Q37D),
      trust_rulingparty  = recode_ab_trust(Q37E),
      trust_opposition   = recode_ab_trust(Q37F),
      trust_police       = recode_ab_trust(Q37G),
      trust_army         = recode_ab_trust(Q37H),
      trust_courts       = recode_ab_trust(Q37I),
      trust_traditional  = recode_ab_trust(Q37J),
      trust_religious    = recode_ab_trust(Q37K),
      age = as.numeric(Q1),
      gender = as.numeric(Q101),
      ethnicity = as.character(Q84A),
      ethnic_unfair = as.numeric(Q88A),
      religion = as.character(Q98A),
      lon = as.numeric(EA_GPS_LO),
      lat = as.numeric(EA_GPS_LA)
    ) %>%
    select(
      round, survey_year, lon, lat, age, gender,
      ethnicity, ethnic_unfair, religion,
      starts_with("trust_")
    )
}
```
#population stuff
```{r}
grid_ea$pop_density <- exactextractr::exact_extract(
  pop_ea,
  grid_ea,
  "mean"
)
#merging to panel
grid_pop <- grid_ea %>%
  st_drop_geometry() %>%
  select(cell_id, pop_density)

panel <- panel %>%
  left_join(grid_pop, by = "cell_id")

#loging density
panel <- panel %>%
  mutate(
    log_pop_density = log1p(pop_density)
  )

```

#stacking rounds
```{r}
ab_all <- bind_rows(
  prep_r6(r6),
  prep_r7(r7),
  prep_r8(r8),
  prep_r9(r9)
) %>%
  filter(
    !is.na(lat), !is.na(lon),
    between(lat, -90, 90),
    between(lon, -180, 180)
  )
```

#building trust index
```{r}
trust_vars <- c(
  "trust_parliament",
  "trust_electoral",
  "trust_localgov",
  "trust_opposition",
  "trust_police",
  "trust_army",
  "trust_courts",
  "trust_traditional",
  "trust_religious"
)

ab_all <- ab_all %>%
  mutate(
    trust_n_nonmiss = rowSums(!is.na(across(all_of(trust_vars)))),
    trust_index_raw = rowMeans(across(all_of(trust_vars)), na.rm = TRUE),
    trust_index_raw = ifelse(trust_n_nonmiss >= 5, trust_index_raw, NA_real_)
  ) %>%
  group_by(round) %>%
  mutate(
    trust_index_z = as.numeric(scale(trust_index_raw))
  ) %>%
  ungroup()
```
#converting to SF and assigning cells
```{r}
ab_sf <- st_as_sf(ab_all, coords = c("lon", "lat"), crs = 4326, remove = FALSE) %>%
  st_make_valid()

ab_sf_ea <- st_transform(ab_sf, crs_ea)

ab_with_cells <- st_join(ab_sf_ea, grid_ea["cell_id"], left = FALSE)
```


#aggregating to cell x round
```{r}
trust_cell_round <- ab_with_cells %>%
  st_drop_geometry() %>%
  group_by(cell_id, round, survey_year) %>%
  summarise(
    trust_cell = mean(trust_index_z, na.rm = TRUE),
    trust_sd   = sd(trust_index_z, na.rm = TRUE),
    n_resp     = n(),
    mean_age   = mean(age, na.rm = TRUE),
    prop_female = mean(gender == 2, na.rm = TRUE),  # adjust coding if needed
    .groups = "drop"
  ) %>%
  filter(n_resp >= 5)
```
#ACLED to cell
```{r}
acled_ea <- st_transform(acled_wa, crs_ea)

acled_with_cells <- st_join(acled_ea, grid_ea["cell_id"], left = FALSE)
#making sure event_date is correct
acled_with_cells <- acled_with_cells %>%
  mutate(
    event_date = suppressWarnings(lubridate::ymd(event_date)),
    event_year = year(event_date)
  )
```

#matching conflict to round
```{r}
round_year_lookup <- tibble(
  round = c(6, 7, 8, 9),
  survey_year = c(2016, 2020, 2022, 2023)
)
#counting conflict to cell x survey_year
conflict_cell_round <- acled_with_cells %>%
  st_drop_geometry() %>%
  inner_join(round_year_lookup, by = c("event_year" = "survey_year")) %>%
  group_by(cell_id, round, event_year) %>%
  summarise(
    conflict_events = n(),
    conflict_any = as.integer(n() > 0),
    .groups = "drop"
  ) %>%
  rename(survey_year = event_year)
```

#merging to analysis data
```{r}
panel <- trust_cell_round %>%
  left_join(conflict_cell_round, by = c("cell_id", "round", "survey_year")) %>%
  mutate(
    conflict_events = replace_na(conflict_events, 0),
    conflict_any    = replace_na(conflict_any, 0)
  )
```
#CFEs

```{r}
grid_country <- st_join(
  st_centroid(grid),
  west_af %>% select(admin = admin),
  left = FALSE
) %>%
  st_drop_geometry() %>%
  select(cell_id, admin)

panel <- panel %>%
  left_join(grid_country, by = "cell_id")

```

#does trust predict conflict? with population
```{r}
m_logit_pop <- feglm(
  conflict_any ~ trust_cell + log_pop_density + n_resp + mean_age + prop_female | admin + round,
  data = panel,
  family = binomial()
)

summary(m_logit_pop)

```

# log odds
```{r}
newdat <- data.frame(
  trust_cell = seq(min(panel$trust_cell, na.rm = TRUE),
                   max(panel$trust_cell, na.rm = TRUE),
                   length.out = 100),
  n_resp = median(panel$n_resp, na.rm = TRUE),
  mean_age = median(panel$mean_age, na.rm = TRUE),
  prop_female = median(panel$prop_female, na.rm = TRUE),
  admin = names(sort(table(panel$admin), decreasing = TRUE))[1],
  round = 9
)

newdat$pred <- predict(m_logit, newdata = newdat, type = "response")

ggplot(newdat, aes(x = trust_cell, y = pred)) +
  geom_line(linewidth = 1) +
  theme_minimal() +
  labs(
    x = "Grid-cell trust",
    y = "Predicted probability of any conflict",
    title = "Predicted conflict probability by trust"
  )
```

#does trust predict number of conflict events? 
```{r}
m_pois_pop <- fepois(
  conflict_events ~ trust_cell + log_pop_density + n_resp + mean_age + prop_female | admin + round,
  data = panel
)

summary(m_pois_pop)
```
